Coalition Formation Algorithm of Prosumers in a Smart Grid Environment

### 2015-02-02 Article.tex
modifs

There seems to be a clear consensus on the benefit of having bidirectional communication flows between the prosumers and the grid, as demand side management relies on such an architecture \cite{WuHamedHuangBook2011}. Nevertheless as the number of active prosumers is expected to rise, it is safe to assume there is a need for more complex communication patterns (prosumers grouping together into so-called "coalitions") that will help to decrease the communication burden and satisfy the multiple requirements of the power grid management. Formation of coalitions inside a smart grid environment could be applied to various types of agents such as self sufficient micro-grids \cite{Pahwa}, sizable and adjustable virtual power plants (VPP) \cite{Braun, Ramchurn}, or fleet of electric vehicles that back up the grid in emergency situations (V2G) \cite{Ramchurn}. These are only a few use cases where the coalition of multiple agents can enhance the grid reliability and in the mean time reduce the communication burden.

In this paper, we consider the case where prosumers wanted to be clustered together into coalitions in order to be able to produce enough power which may be sold to the grid. If the coalition's production is sufficiently high and stable, the coalition can negociate on a day by day basis with the grid by announcing each day an expected production for the next day. We focus on how to statistically improve the production stability by carefully forming coalitions of prosumers that have greater probabilities of staying in acceptable range of production. By using prediction techniques, we form coalitions that meet the contract values ranges proposed by the grid, enabling it to schedule its production accordingly. Meeting contract values for the prosumers is especially relevant in power trade conditions, where energy is traded based on day-ahead predictions. In this context, participants should try to minimize their prediction errors in order to maintain a stable state for the grid. They may even occur some penalties if the productions deviate from the initial contract values. However, it remains pertinent for both side to maintain the production as stable as possible : for the prosumer, the stability (with renewable energy sources) will impact its net gain, and for the power grid, it will help to maintain its reliability.

@@

In this paper we do not consider stability against player defection but focus on a statistics-oriented definition of coalition stability. We want to form coalitions based on utilities that depend on statistical properties of time-series values. The setting is thus similar with some financial studies on stock exchanges, where researchers tried to find relevant correlated clusters of stocks based on their daily prices variations. There exists well-known algorithms such as K-means or hierarchical clustering that are traditionally used for such purposes. Nevertheless, despite their popularity, they do not appear particularly well suited for meaningfully translating complex correlation relationships between time series while clustering them, especially when these series exhibit complicated underlying patterns. We will show later in this paper that using K-means as a clustering technique for grouping prosumers together leads to less good results. In \cite{Mantegna1999} the author introduced an approach where stocks time-series of their daily log returns, are organized in a graph such that stocks exhibiting similar price fluctuation patterns are close to each other. This closeness notion is formalized with a similarity measure based on Pearson correlation coefficient (see notations in table \ref{table1}) ($ d_{ij} = \dfrac{1}{2}\sqrt{2(1-\rho_{ij})} $ or $ d_{ij} = 1 - \rho_{ij}^{2} $ ) that enables to weight the edges of the graph.

@@

Presented in this way, the time-series clustering task seems very close to graph community detection. Communities in networks are indeed often seen as groups of nodes exhibiting high internal densities of links as well as a low density across communities, and several topology oriented techniques for finding communities are present in the litterature (\cite{Newman2013} \cite{Girvan2002} \cite{Newman2013_2}). For our purpose where decorrelation is the closeness notion, such algorithms tend to have some difficulties because only a few inclusions of very high correlations can strongly affect the stability of a coalition. Furthermore, these graphs tend to exhibit far less clustering than their correlation-based counterparts, which can increase the difficulty of finding communities. However, we will see in the following, that a clique (a complete subgraph) in our filtered decoralation-based graph exhibit, by construction, low correlation. We indeed removed in the filtering (see $\epsilon$-graphs) all links that stands for high correlation. There exists greedy heuristics based on cliques in the litterature such as clique percolation \cite{Lancichinetti} which realizes local optimizations of a fitness function and results in overlaping community structures. Detection of overlapping communities is actually a very active field of research, especially in social networks where a person might belong to several communities. Section \ref{sec:forming} will explain how we used such techniques on the decorelation graph in order to form coalitions of prosumers.

@@

Our major concern is to characterize a prosumer behavior with realistic patterns of consumption, production as well as realistic geographical correlations between them. Only wind turbines and PV (photovoltaic panels) will be used as generators throughout this paper, and we denote by $ \mathcal{A} = \{ a_{1},...a_{N} \} $ the set of prosumers. To account for the production of a prosumer, we rely on French's weather data from 2006 to 2012 sampled every 3 hours \cite{Infoclimat} (similar data are available for the United States covering 2010\cite{NCDC}). As shown in the first part of figure 1, the first step consists in discretizing the studied zone around well chosen weather stations and gathering three kinds of traces (wind speed, cloudiness, and temperature) that we organize in a structure we will refer to as "climate vector" in the following. We consider that climate vectors are constant over their area, meaning that, if two agents are in the same area, they are exposed to the same climate vector. We use existing DER models enabling conversion between weather variables and output powers (see power curves in \cite{Kota2011} \cite{windturbinemodel}). \cite{Dans2007} provides also a convenient way of using cloudiness traces as realistic degradation factors in a \textit{"clear blue sky"} solar model, which enables us to rebuild realistic solar irradiance traces for stations that do not provide such information directly.

@@

In the following, we do not make any gaussian assumption as in the previous toy example. We investigate then the developement of a heuristic that tries to find a compromise between maximizing the production sold on the market (benefits) on one hand and minimizing the probability that the coalitions fail to sustain their contract values (stability) on the other hand (Table \ref{table1} summarizes the main notations) :

@@

\begin{table}[t]
\centering
\scriptsize
\begin{tabular}{l|p{4.5cm}}
  \hline 
  \\
  $ d_{ij} $ & Distance metric between timeseries i and j \\
  $ \rho_{ij} $ & Pearson's correlation coefficient between timeseries i and j \\
  $ \mathcal{A} $ & Agent set \\
  $ P_{i}(t) \in \mathbb{R} $ & Instantaneous production of agent $i$ at time $t$ \\
  $ P_{s}(t) $ & Instantaneous production of coalition s \\
  $ \mathcal{F}(P_{i}) $ & Probability distribution of $i$'s production \\
  $ \mathcal{F}(P_{S}) $ & Probability distribution of coalition s production \\
  $ \mathcal{N}(\mu, \sigma) $ & Normal distribution with mean $ \mu $ and standard deviation $ \sigma $ \\
  $ \phi \in [0,1] $ & Reliability threshold \\
  $ P^{MIN} \in \mathbb{R}^+ $ & Min value threshold \\
  $ P_{s}^{CRCT} \in \mathbb{R}^+ $ & Contract value announced by coalition s \\
  $ P_{\phi}(s) \in \mathbb{R}^+ $ & Max contract value under $ \phi $ constraint \\
  $ \mathcal{U}_{\phi,\ P^{MIN}}(s) $ & Utility function \\
  $ \tau_{i}(s) \in \mathbb{R}$ & Overlapping to disjoint coalitions mapping \\
  $ \epsilon \in [0,1] $ & Pruning threshold for the correlation graph \\
  $ G_{\epsilon}(\mathcal{A}, E) $ & Filtered correlation graph \\
  $ N_{COAL} \in \mathbb{Z}^+ $ & Number of desired coalition \\
  $ CS $ & Set of all coalition structures \\
  $ \Theta(G) $ & Set of non overlaping cliques in graph G \\
  \\
  \hline
\end{tabular}

@@

However, these clusters of strongly correlated time-series are the exact opposite of what we are seeking. We can indeed consider them directly as coalitions and compute their utilities, and the results show (see the green curves with squares on figure \ref{Fig4}), as expected, terrible values. We will indeed consider in section \ref{sec:results} this case as a worst case scenario, far worse than a trivial random split of the agents in the same number of coalitions.

@@

However, as seems intuitively understandable, cliques of this graph tend to exhibit very good utility values. Such  structures contain indeed a link between every two nodes, meaning that the overall correlation is quite small. Obviously, the $ \epsilon $ parameter is indirectly responsible for the sizes of the cliques : if it is too low, $ G_{\epsilon} $ does not provide enough cliques, conversely, if $\epsilon $ is too high, we loose important information as the graph becomes very dense and cliques tend to overlap strongly. For large values of $ \epsilon $ and for non trivial number of agents, finding cliques can even become computationally intensive. Despite being direct and simple, improving the sizes of the cliques by increasing $ \epsilon $ does not seem elegant. Furthermore, as the utility function is also focused on maximizing $ P_{\phi}(s) $, it might be the case that some cliques benefit from additional agents even if they don't form a clique anymore. Cliques for small values of $ \epsilon $ appear thus as good seeds for forming stable coalitions rather than directly acceptable coalitions. We define $ N_{COAL} $ as the desired number of coalitions fixed by the user. Furthermore, let $ \Theta(G_{\epsilon}) $ be the set of non overlaping cliques in $ G_{\epsilon} $, we choose the $ \epsilon $ parameter as the smallest value such that $ G_{\epsilon} $ contains at least $ N_{COAL} $ non overlaping cliques :

@@

We construct $ G_{\epsilon^{\star}} $ such that it contains at least $ N_{COAL} $ potential seeds that may or may not be above the grid requirements, but exhibit low correlations among the members. Naturally, grid requirements could eventually be reached and social welfare increased by incorporating more nodes into the seeds. For this purpose, we use a clique percolation algorithm on $ G_{\epsilon^{\star}} $ that will extend the seeds over a correlation constrained environment. The main idea behind this algorithm is to make the seeds grow by local optimization of a fitness function, usually based on internal and external degrees of the seeds \cite{Lancichinetti}. The main difference between our case study and this algorithm for community detection is that we chose the fitness function as our utility function (cf. eq.\eqref{utility}) that both maximizes the production and minimizes its volatility (see eq.\ref{problem_1} and eq.\ref{problem_2}). As explained in section \ref{sec:related}, clique percolation leads generally to overlapping communities. For simplicity, in this paper, we wish to keep the coalitions disjoint and  leave the management of overlapping coalitions for future work. We thus implemented a simple heuristic that considers nodes in multiple seeds one by one and chooses its final coalition as the one that “needs it the most” in terms of utility loss. More formally, for a coalition \textit{s} and a node $ i \in s $, we define\footnote{\label{note1}Not to be mingled with the Shapley value concept} :

@@

As shown in figure \ref{Fig2}a, the parameters $\phi$ and $P^{MIN}$ shape the utility function such that, if $ \phi $ is close to zero, the reliability requirement is very high. Only small values of $ P^{MIN}$ could then potentially lead to valid coalitions (and positive utilities). Conversely, the higher $\phi$, the less constraints are imposed to the coalitions and more valid coalitions can arise for a larger spectrum of $ P^{MIN}$. Obviously, the highest utility values are found for high $ \phi $, because the formed coalitions are able to announce higher contract values, yielding higher utilities.

@@

In more details, figure \ref{Fig4}a presents the evolution of social welfare as the number of coalitions increases (all other parameters are kept constant) for random (red curve with squares), clique percolation (blue curve with circles), and correlated coalitions (green curves with triangles) that stands for a worst case scenario. As for figure \ref{Fig4}b, it shows the percentage of coalitions able to enter the market for different values of $ P^{MIN} $. For consistency, we average the results of both plots over 100 realizations (colored zones around the curves shows the standard deviations). When the grid requirements are constant (figure \ref{Fig4}b), and the number of desired coalitions is low, clique percolation generally performs better than random case. Moreover, when $ N_{COAL} $ gets bigger, the performance of a random split tumble down rapidly while our clique percolation is able to maintain efficiently the social welfare of the coalitions formed. When the grid requirements vary, for very low $ P^{MIN} $, all coalitions for all algorithms are able to enter the market, yielding an acceptance percentage of $ 100 \% $. But as $ P^{MIN} $ increases, we see the percentage of the correlated coalitions is going down very quickly. After a few increases in $ P^{MIN} $, the percentage of the random procedure starts dropping while it stays almost constant for our algorithm. For $ P^{MIN} = 8 $, we see that only a little more than half of the coalitions for the random case are able to enter while approximately $ 85 \%$ of them enters for the clique percolation algorithm. Finally, when the grid requirements becomes too high, the acceptance percentage of our algorithm tends slowly to zero (not shown in this plot for readability).

@@

We believe that the originality of this paper lies in its willingness to exploit the de-correlation of prosumer profiles in order to build stable coalitions. In this direction, we presented a model based on meteorological traces, that captures the complex "\textit{energy mix/climate vectors}" combination and generates realistic production and consumption patterns. We then built a framework that enables the grid to specify stability and minimum production requirements for filtering the coalitions. On this basis, we proposed a simple algorithm that seeks for uncorrelated prosumer patterns as potential seeds and expand them in coalitions able to rise above the grid requirements. We validated our algorithm against a random choice of coalitions, which can be considered as an average case. We showed  that it performs better (coalitions are more stable and the overall production is more important) and that it exhibits a higher robustness/flexibility against grid requirements changes. The worst case scenario is represented by coalitions of correlated prosumers. Interesting leads for future works would be the use of correlated clusters to reduce the number of entities the algorithm has to deal with, or the introduction of a payoff allocation towards the prosumers such that the stability against player defection could be analyzed through game theory. Showing that maintaining the coalitions formed with our algorithm necessitates less communication and less storage capacity could also lead to a stimulating project. Besides, we believe that not restricting the algorithm to non overlapping coalitions and studying the strategies and weights of nodes with multiple options could lead to interesting research.

>>>>

There seems to be a clear consensus on the benefit of having bidirectional communication flows between the prosumers and the grid, as demand side management relies on such an architecture \cite{WuHamedHuangBook2011}. Nevertheless as the number of active prosumers is expected to rise, it is safe to assume there is a need for more complex communication patterns (prosumers grouping together into so-called "coalitions") that will help to decrease the communication burden and satisfy the multiple requirements of the power grid management.

Formation of coalitions inside a smart grid environment could be applied to various types of agents such as self sufficient micro-grids \cite{Pahwa}, sizable and adjustable virtual power plants (VPP) \cite{Braun, Ramchurn}, or fleet of electric vehicles that back up the grid in emergency situations (V2G) \cite{Ramchurn}. These are only a few use cases where the coalition of multiple agents can enhance the grid reliability and in the mean time reduce the communication burden.

In this paper, we consider the case where prosumers wanted to be clustered together into coalitions in order to be able to produce enough power which may be sold to the grid. If the coalition's production is sufficiently high and stable, the coalition can negociate on a day by day basis with the grid by announcing each day an expected production for the next day.

We focus on how to statistically improve the production stability by carefully forming coalitions of prosumers that have greater probabilities of staying in acceptable range of production. By using prediction techniques, we form coalitions that meet the contract values ranges proposed by the grid, enabling it to schedule its production accordingly. Meeting contract values for the prosumers is especially relevant in power trade conditions, where energy is traded based on day-ahead predictions.

In this context, participants should try to minimize their prediction errors in order to maintain a stable state for the grid. They may even occur some penalties if the productions deviate from the initial contract values. However, it remains pertinent for both side to maintain the production as stable as possible : for the prosumer, the stability (with renewable energy sources) will impact its net gain, and for the power grid, it will help to maintain its reliability.

@@

In this paper we do not consider stability against player defection but focus on a statistics-oriented definition of coalition stability. We want to form coalitions based on utilities that depend on statistical properties of time-series values. The setting is thus similar with some financial studies on stock exchanges, where researchers tried to find relevant correlated clusters of stocks based on their daily prices variations. There exists well-known algorithms such as K-means or hierarchical clustering that are traditionally used for such purposes. Nevertheless, despite their popularity, they do not appear particularly well suited for meaningfully translating complex correlation relationships between time series while clustering them, especially when these series exhibit complicated underlying patterns. We will show later in this paper that using K-means as a clustering technique for grouping prosumers together leads to less good results.

In \cite{Mantegna1999} the author introduced an approach where stocks time-series of their daily log returns, are organized in a graph such that stocks exhibiting similar price fluctuation patterns are close to each other. This closeness notion is formalized with a similarity measure based on Pearson correlation coefficient (see notations in table \ref{table1}) ($ d_{ij} = \dfrac{1}{2}\sqrt{2(1-\rho_{ij})} $ or $ d_{ij} = 1 - \rho_{ij}^{2} $ ) that enables to weight the edges of the graph.

@@

Presented in this way, the time-series clustering task seems very close to graph community detection. Communities in networks are indeed often seen as groups of nodes exhibiting high internal densities of links as well as a low density across communities, and several topology oriented techniques for finding communities are present in the litterature (\cite{Newman2013} \cite{Girvan2002} \cite{Newman2013_2}).

For our purpose where decorrelation is the closeness notion, such algorithms tend to have some difficulties because only a few inclusions of very high correlations can strongly affect the stability of a coalition. Furthermore, these graphs tend to exhibit far less clustering than their correlation-based counterparts, which can increase the difficulty of finding communities.

However, we will see in the following, that a clique (a complete subgraph) in our filtered decoralation-based graph exhibit, by construction, low correlation. We indeed removed in the filtering (see $\epsilon$-graphs) all links that stands for high correlation. There exists greedy heuristics based on cliques in the litterature such as clique percolation \cite{Lancichinetti} which realizes local optimizations of a fitness function and results in overlaping community structures.

Detection of overlapping communities is actually a very active field of research, especially in social networks where a person might belong to several communities. Section \ref{sec:forming} will explain how we used such techniques on the decorelation graph in order to form coalitions of prosumers.

@@

Our major concern is to characterize a prosumer behavior with realistic patterns of consumption, production as well as realistic geographical correlations between them. Only wind turbines and PV (photovoltaic panels) will be used as generators throughout this paper, and we denote by $ \mathcal{A} = \{ a_{1},...a_{N} \} $ the set of prosumers. To account for the production of a prosumer, we rely on French's weather data from 2006 to 2012 sampled every 3 hours \cite{Infoclimat} (similar data are available for the United States covering 2010\cite{NCDC}).

As shown in the first part of figure 1, the first step consists in discretizing the studied zone around well chosen weather stations and gathering three kinds of traces (wind speed, cloudiness, and temperature) that we organize in a structure we will refer to as "climate vector" in the following. We consider that climate vectors are constant over their area, meaning that, if two agents are in the same area, they are exposed to the same climate vector. We use existing DER models enabling conversion between weather variables and output powers (see power curves in \cite{Kota2011} \cite{windturbinemodel}). \cite{Dans2007} provides also a convenient way of using cloudiness traces as realistic degradation factors in a \textit{"clear blue sky"} solar model, which enables us to rebuild realistic solar irradiance traces for stations that do not provide such information directly.

@@

In the following, we do not make any gaussian assumption as in the previous toy example. We investigate then the developement of a heuristic that tries to find a compromise between maximizing the production sold on the market (benefits) on one hand (see eq. \ref{problem_1}) and minimizing the probability that the coalitions fail to sustain their contract values (stability) on the other hand (see eq. \ref{problem_2}). Table \ref{table1} summarizes the main notations.

@@

\begin{table}[t]
\centering
\scriptsize
\begin{tabular}{l|p{4.5cm}}
  \hline 
  \\
  $ d_{ij} $ & Distance metric between timeseries i and j \\
  $ \rho_{ij} $ & Pearson's correlation coefficient between timeseries i and j \\
  $ \mathcal{A} $ & Agent set \\
  $ P_{i}(t) \in \mathbb{R} $ & Instantaneous production of agent $i$ at time $t$ \\
  $ P_{s}(t) $ & Instantaneous production of coalition s \\
  $ \mathcal{F}(P_{i}) $ & Probability distribution of $i$'s production \\
  $ \mathcal{F}(P_{S}) $ & Probability distribution of coalition s production \\
  $ \mathcal{N}(\mu, \sigma) $ & Normal distribution with mean $ \mu $ and standard deviation $ \sigma $ \\
  $ \phi \in [0,1] $ & Reliability threshold \\
  $ P^{MIN} \in \mathbb{R}^+ $ & Min value threshold \\
  $ P_{s}^{CRCT} \in \mathbb{R}^+ $ & Contract value announced by coalition s \\
  $ P_{\phi}(s) \in \mathbb{R}^+ $ & Max contract value under $ \phi $ constraint \\
  $ \mathcal{U}_{\phi,\ P^{MIN}}(s) $ & Utility function \\
  $ \tau_{i}(s) \in \mathbb{R}$ & Overlapping to disjoint coalitions mapping \\
  $ \epsilon \in [0,1] $ & Pruning threshold for the correlation graph \\
  $ G_{\epsilon}(\mathcal{A}, E) $ & Filtered correlation graph \\
  $ N_{COAL} \in \mathbb{Z}^+ $ & Number of desired coalition \\
  $ CS $ & Set of all coalition structures \\
  $ \Theta(G) $ & Set of non overlaping cliques in graph G \\
  $ \mathbf{1}_{X} $ & Indicator function over set X \\ 
  \\
  \hline
\end{tabular}

@@

However, these clusters of strongly correlated time-series are the exact opposite of what we are seeking. We can indeed consider them directly as coalitions and compute their utilities, and the results show (see the green curves with squares on figure \ref{Fig4}), as expected, terrible values. We will indeed consider in section \ref{sec:results} this case as a worst case scenario, far worse than a trivial random split of the agents in the same number of coalitions (see the red curves with triangles of the random case on figure \ref{Fig4}).

@@

However, as seems intuitively understandable, cliques of this graph tend to exhibit very good utility values. Such  structures contain indeed a link between every two nodes, meaning that the overall correlation is quite small.

Obviously, the $ \epsilon $ parameter is indirectly responsible for the sizes of the cliques : if it is too low, $ G_{\epsilon} $ does not provide enough cliques, conversely, if $\epsilon $ is too high, we loose important information as the graph becomes very dense and cliques tend to overlap strongly. For large values of $ \epsilon $ and for non trivial number of agents, finding cliques can even become computationally intensive.

Despite being direct and simple, improving the sizes of the cliques by increasing $ \epsilon $ does not seem elegant. Furthermore, as the utility function is also focused on maximizing $ P_{\phi}(s) $, it might be the case that some cliques benefit from additional agents even if they don't form a clique anymore. Cliques for small values of $ \epsilon $ appear thus as good seeds for forming stable coalitions rather than directly acceptable coalitions.

We define $ N_{COAL} $ as the desired number of coalitions fixed by the user. Furthermore, let $ \Theta(G_{\epsilon}) $ be the set of non overlaping cliques in $ G_{\epsilon} $, we choose the $ \epsilon $ parameter as the smallest value such that $ G_{\epsilon} $ contains at least $ N_{COAL} $ non overlaping cliques :

@@

We construct $ G_{\epsilon^{\star}} $ such that it contains at least $ N_{COAL} $ potential seeds that may or may not be above the grid requirements, but exhibit low correlations among the members. Naturally, grid requirements could eventually be reached and social welfare increased by incorporating more nodes into the seeds. For this purpose, we use a clique percolation algorithm on $ G_{\epsilon^{\star}} $ that will extend the seeds over a correlation constrained environment. The main idea behind this algorithm is to make the seeds grow by local optimization of a fitness function, usually based on internal and external degrees of the seeds \cite{Lancichinetti}. The main difference between our case study and this algorithm for community detection is that we chose the fitness function as our utility function (cf. eq.\eqref{utility}) that both maximizes the production and minimizes its volatility (see eq.\ref{problem_1} and eq.\ref{problem_2}).

As explained in section \ref{sec:related}, clique percolation leads generally to overlapping communities. For simplicity, in this paper, we wish to keep the coalitions disjoint and  leave the management of overlapping coalitions for future work. We thus implemented a simple heuristic that considers nodes in multiple seeds one by one and chooses its final coalition as the one that “needs it the most” in terms of utility loss. More formally, for a coalition \textit{s} and a node $ i \in s $, we define\footnote{\label{note1}Not to be mingled with the Shapley value concept} :

@@

As shown in figure \ref{Fig2}a, the parameters $\phi$ and $P^{MIN}$ shape the utility function such that, if $ \phi $ is close to zero, the reliability requirement is very high. Only small values of $ P^{MIN}$ could then potentially lead to valid coalitions (and positive utilities). Conversely, the higher $\phi$, the less constraints are imposed to the coalitions and more valid coalitions could be formed for a larger spectrum of $ P^{MIN}$. Obviously, the highest utility values are found for high $ \phi $, because the formed coalitions are able to announce higher contract values, yielding higher utilities.

@@

In more details, figure \ref{Fig4}a presents the evolution of social welfare as the number of coalitions increases (all other parameters are kept constant) for random (red curve with triangles), clique percolation (blue curve with circles), and correlated coalitions (green curves with squares) that stands for a worst case scenario.

As for figure \ref{Fig4}b, it shows the percentage of coalitions able to enter the market for different values of $ P^{MIN} $. For consistency, we average the results of both plots over 100 realizations (colored zones around the curves shows the standard deviations). When the grid requirements are constant (figure \ref{Fig4}b), and the number of desired coalitions is low, clique percolation generally performs better than random case. Moreover, when $ N_{COAL} $ gets bigger, the performance of a random split tumbles down rapidly while our clique percolation is able to maintain efficiently the social welfare of the coalitions formed. When the grid requirements vary, for very low $ P^{MIN} $, all coalitions for all algorithms are able to enter the market, yielding an acceptance percentage of $ 100 \% $. But as $ P^{MIN} $ increases, we see the percentage of the correlated coalitions is going down very quickly.

After a few increases in $ P^{MIN} $, the percentage of the random procedure starts dropping while it stays almost constant for our algorithm. For $ P^{MIN} = 8 $, we see that only a little more than half of the coalitions for the random case are able to enter while approximately $ 85 \%$ of them enters for the clique percolation algorithm. Finally, when the grid requirements become too high, the acceptance percentage of our algorithm tends slowly to zero (not shown in this plot for readability).

@@

We believe that the originality of this paper lies in its willingness to exploit the de-correlation of prosumer profiles in order to build stable coalitions. In this direction, we presented a model based on meteorological traces, that captures the complex "\textit{energy mix/climate vectors}" combination and generates realistic production and consumption patterns. We then built a framework that enables the grid to specify stability and minimum production requirements for filtering the coalitions.

On this basis, we proposed a simple algorithm that seeks for uncorrelated prosumer patterns as potential seeds and expand them in coalitions able to rise above the grid requirements. We validated our algorithm against a random choice of coalitions, which can be considered as an average case. We showed  that it performs better (coalitions are more stable and the overall production is more important) and that it exhibits a higher robustness/flexibility against grid requirements changes. The worst case scenario is represented by coalitions of correlated prosumers.

Interesting leads for future works would be the use of correlated clusters to reduce the number of entities the algorithm has to deal with, or the introduction of a payoff allocation towards the prosumers such that the stability against player defection could be analyzed through game theory. Showing that maintaining the coalitions formed with our algorithm necessitates less communication and less storage capacity could also lead to a stimulating project. Besides, we believe that not restricting the algorithm to non overlapping coalitions and studying the strategies and weights of nodes with multiple options could lead to interesting research.
